Job mobility and earnings growth: diffrerent markets, different outcomes.

### 2018-10-15 draft.Rmd
Update draft.Rmd

**Design/methodology/approach:** We present parallel case studies using the British Household Panel Survey, and the German Socio-Economic Panel. Within both panels, we consider job history questions which operationalise voluntary internal and external job mobility. Using this data we estimate the effect of job mobility on earnings growth using fixed-effects estimation, and compare both countries.

**Findings:** Internal and external mobility are distinctly different strategies for improving earnings. However, the institutional setting under which mobility takes place also shapes the rewards tied to mobility. In both countries the core workforce benefits from internal promotions. However, in Germany the rewards tied to promotion are far greater than they are in the UK. A key reason German workers are less mobile than British workers may have less to do with employment protection legislation and more to do with the rewards tied to internal career ladders.

@@

**Originality/value:** Studies exploring the effects of job mobility tend to focus on single country cases. The lessons from such cases are then prescribed to other countries, without acknowledging that countries differ by capitalisms and institutional settings. Our paper articulates important differences in the mobility-earnings relationship between Germany and the UK, and helps to understand processes of inequality in both countries. German workers who are able to pursue internal promotion are likely different form those who are unable to pursue promotion. British workers also benefit from promotions, but must pursue several promotions before seeing a substantial difference in pay.

@@

Voluntary job mobility is often seen as an important way to improve individual earnings and conditions. Routinely, authors show how workers make significant gains through quitting jobs or securing promotions. However, capitalist countries differ in mobility rates; especially regarding rates of voluntary mobility. If quitting a job is inherrently beneficial to workers, but more common in the UK than Germany, does this mean that German workers are held back from potential gains, or does it mean that the nature of quitting in both countries is different? Authors typically rationalise such differences as a difference in employment protection legislation (Brown) or a difference in job creation and destruction (OECD). This article takes a different approach, and shows that German and British institutions shape not only the likelihood of job mobility, but also the rewards tied to mobility.

>>>>

**Design/methodology/approach:** We consider parallel case studies using the British Household Panel Survey, and the German Socio-Economic Panel. In both panels, we use job history questions which operationalise voluntary internal and external job mobility. Using this data we estimate the effect of job mobility on earnings growth using fixed-effects estimation, and compare both countries.

**Findings:** Internal and external mobility are distinctly different strategies for improving earnings. However, the institutional setting, where mobility types take place, also shapes mobility's impact on earnings growth. In both countries the core workforce benefits from internal promotions. However, in Germany the rewards tied to promotion are far greater than they are in the UK. A key reason German workers are less mobile than British workers may have less to do with employment protection legislation and more to do with the rewards tied to internal career ladders.

@@

**Originality/value:** The majority of studies which explore the effects of mobility focus on single country cases. The lessons from such cases are then prescribed to other countries, without acknowledging institutional differences and settings. Our paper such settings in the mobility-earnings relationship between Germany and the UK, and helps to understand processes of inequality in both countries. German workers who are able to pursue internal promotion are likely different from those who are unable to pursue promotion. British workers also benefit from promotions, but must pursue several promotions before seeing a substantial difference in pay.

@@

Researchers see the process of voluntary job mobility as crucial to understanding inequality. Quits and promotions, therefore, are important strategies which workers use to improve individual earnings and conditions. For this reason, many sociologists study worker access to quits and promotions, and the class and gender differences behind such rates. Further, both economists and sociologists show how workers manage to secure significant gains through quitting jobs or securing promotions. However, capitalist countries differ in mobility rates; especially regarding rates of voluntary mobility. If quitting a job is inherrently beneficial to workers, but more common in the UK than Germany, does this mean that German workers are held back from potential gains, or does it mean that the nature of quitting in both countries is different? Authors typically rationalise such differences as a difference in employment protection legislation (Brown) or a difference in job creation and destruction (OECD). This article takes a different approach, and shows that German and British institutions shape not only the likelihood of job mobility, but also the rewards tied to mobility.
